Three Problems for Sophistication

### 2015-01-19 Facticity/Facticity/paper.tex

Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring. There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity. For example, a sequence of a million coin flips will almost certainly have maximal Kolmogoroc complexity, even though there is nothing complex about flipping a lot of coins---it's just hard work. For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data. This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed. In this paper, we investigate three serious problems with these definitions. We believe that ultimately, some of these issues are insurmountable, and provide several intuitive arguments why any attempt to define sophistication must fail.

The Kolmogorov complexity of a binary string $x$ is, roughly, the length of the shortest computer program that will output $x$. This length depends on which programming language is used, but only by a constant independent of $x$. This property is known as \emph{invariance}. This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and we can see the Kolmogorov complexity  as \emph{a property of the data}.

A definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees. We are interested in a measure satisfying the following requirements:

@@

Underfitting (Section~\ref{section:underfitting}) occurs when the selected model is simple, but fails to represent some of the structure present in the data. In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language. Such \emph{universal models} can be described in a constant amount of bits, and  represent the data with a program the length of the Kolmogorov complexity, essentially encoding all information as noise. If complex models are penalized, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data. The usual workaround is to restrict the set of allowed models to exclude the universal ones. For example, the model class may be restricted to the set of all total recursive functions. While excludes universal models, it is questionable whether it adequately solves the problem of underfitting in general.

@@

\subsection{Notation}

Let $\B = \{0,1\}^*$. We deal with computable partial functions $\phi: \B \to \B$, which we will call \emph{models}. A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix free set. A function $f$ is \emph{total} if $D_f = \B$. A prefix-free function is \emph{complete} if every infinite binary string has a member of $D_f$ as a prefix. We fix a prefix code with $|\bar{x}| \le|x|+2\log(|x|+2)|$ (eg. \cite[Section~1.4]{li1993introduction}).

\begin{definition}[Numberings]
A numbering is an enumeration of of the partial computable functions, denoted as $\phi_1, \phi_2, \ldots$ or simply $\phi$. We fix one effective numbering $\phi$. We call a numbering $\num{\psi}$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ with $\forall: i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{b(i)}$.
\end{definition}

\begin{definition}[Complexity]
Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions. We define the plain and prefix Kolmogorov complexity as $C(x)=\min\{|\bar\imath y|:\phi_i(y)=x\}$ and $K(x)=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}$ respectively.
\label{definition:complexity}
\end{definition}

\begin{definition}[Model classes] A \emph{model class} is a set of models. We distinguish four important model classes: the partial (computable) functions ($\C$), the prefix functions ($\K \subset \C$), the total functions ($\T \subset \C$), the complete functions ($\P \subset \K$) and the finite sets ($\F \subset \P$, encoded as a prefix-free function mapping all strings of length $\lceil\log S\rceil$ to members of $S$).
\end{definition}

\begin{definition}[Representations]
A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
Define the following code-length functions for the representations:\\
\-\hspace{1cm}$L(m)$ is the length of a self-delimiting description of the model $m$.\\ 
\-\hspace{1cm}$L^m(x)=L(m)+\min\{|y|:m(y)=x\}$\\
\-\hspace{1cm}$L^\M(x)=\min\{L^m(x):m\in\M\}$
\end{definition}
We can now define sophistication generically:
\begin{definition}[Sophistication]
The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
$C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}$
with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $L^\M(x)$. The \emph{sophistication} is \\ 
\-\hspace{1cm}$\s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}$\label{definition:sophistication}
\end{definition}
Finally, we generalize the notion of invariance to allow for less strict versions. 
\begin{definition}[$\gamma-invariance$]
Let $\gamma:\N\to \N$ be some nondecreasing function. Two functions $f$ and $g$ are $\gamma$-equivalent if $|f(x) - g(x)| \in O(\gamma(\max(f(x), g(x)))$. We say that a function $\s_\phi(x)$ is \emph{$\gamma$-invariant} to the choice of numbering if $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$. 
\end{definition}

\begin{lemma}
If a function $\s_\psi(x)$ is bounded for some $\psi$, it is either not $\gamma$-invariant or it is  $\gamma$-equivalent to a bounded function.
\end{lemma}

\section{Inefficient indices}
\label{section:indices}
We start from a simple intuition: the definition of $K$ already uses a description of the data in two parts; a computable function and an input. Can we assume that, when such a description is optimal, it automatically separates the structural and random information? As we will show, the representation of functions plays crucial fundamental.

We will first illustrate what happens if any acceptable numbering is allowed:
\begin{restatable}{lemma}{ineffindices}
Let $\phi$ be some acceptable numbering, and $\s_\phi(x)$ be any function which gives the length of some index $i$ such that $\phi_i(y) = x$ for some $y$.

No such function can be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
\end{restatable}
\noindent In the proof of this lemma (Section~\ref{section:appendix-inefficient-indices}), we deliberately used ridiculously inefficient index functions. We could use a more efficient index, but most definitions of sophistication do not mention any such restrictions. As we shall see, any inefficiency in the index will lead to a lack of invariance in sophistication.

@@

\begin{definition}[\cite{grunwald2004shannon,vitanyi2004meaningful}]
The complexity of a partial recursive function is defined by $C_\phi(f) = \min\{C(i):\phi_i=f\}$ and $K_\phi(f) = \min\{K(i):\phi_i=f\}$.\label{definition:model-complexity}
\end{definition}
We show in the appendix (Lemma~\ref{lemma:invariance}) that these are invariant, ie. complexity is a property of a function. There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature.
\begin{enumerate}
  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used. Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$. This approach is taken by Adriaans \cite{adriaans2012facticity}.
  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $K$ to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data. With this approach, the total code length is $K(\phi_i)+|y|$, with $\phi_i(y)=x$. In this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints. This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
\end{enumerate}
To show that the two perspectives are equivalent, we must show that there exists an acceptable numbering which is efficient enough:
\begin{definition}[Faithful Numberings]\label{def:faithful}
  A numbering $\num{\psi}$ is \emph{faithful} if
  for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and
  $|j|\le C(\psi_j)+c$, for some constant $c$.
\end{definition}
This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}. We show in the appendix (Lemma~\ref{lemma:faithful-numberings}) that such numberings exist. Unfortunately, even with a faithful numbering, we can fail:
\begin{restatable}{lemma}{ineffprefix}
Let $\phi$ be faithful and let $L(\phi_i) = |\bar\imath|$, for some prefix encoding function. If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $\s^\C$ is bounded for all numberings.\label{lemma:prefix-inefficiency}
\end{restatable}
\noindent This result applies to \cite{adriaans2012facticity}. To avoid the problem, the model $\phi_i$ should be encoded in $K(\phi_i)$ bits, relying on $K$'s own prefix encoding to delimit the model.

Note that the condition for this lemma likely applies to all possible computable  prefix encoding functions (see Section~\ref{section:prefix-inefficiency}). Thus, even if we have a faithful numbering, we can only use it if we don't count the cost of delimiting the model from the input (as in \cite{koppelSoph1988}). In the following, we will take the second perspective: the representations are two-part descriptions $(i, y)$ such that $\phi_i(y)=x$, with codelength $K(\phi_i) + |y|$. 
\section{Overfitting: the singletons}
\label{section:overfitting}
We now come to the main issue plaguing sophistication. If we use model class $\K$, the candidate set always contains two representations that do not provide a good separation. The first is a \emph{universal model}: a basic universal Turing machine. The second is a \emph{singleton model} a model which returns $x$ for any input. As we will show, in many settings, we can ensure that one of them is the \emph{only} model in the candidate set. The following lemma generalizes the principle:

\begin{restatable}{lemma}{coolone}
\label{lemma:thecoolone}
  Let $\psi_1,\psi_2,\ldots$ be any acceptable enumeration of the partial recursive functions.
  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\{0,1\}^*\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e. if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$. Let $\M'=\{\psi_i:i\in\tn{range}(D)\}$. Further assume there is a constant $c$ such that:\\
\-\hspace{1cm}(1)$\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$\\
\-\hspace{1cm}(2)$\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.\\
Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
\end{restatable}
\noindent We can now show numberings for which the singletons are always selected:

\begin{restatable}[Overfitting]{lemma}{overfitting}
Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$. Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
\[
K(x)-S^{\M,\phi}(x)\le c
\]
for all $x\in\X$.
\end{restatable}
\noindent In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication. This means that for any dataset, the sophistication is equal to the Kolmogorov complexity.

Many variants in the literature (including the structure function) are susceptible to this problem \cite{cover1985kolmogorov,gacs2001algorithmic,vereshchagin2013algorithmic}. \cite{vitanyi2004meaningful,adriaans2012facticity} avoid it by not using self-delimiting representations. Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically not optimally efficient. While this dodges the issue of the singletons, it does not show that $S$ is invariant.

@@

\section{Underfitting: the universal model}
\label{section:underfitting}
On the other end of the spectrum, there exist representations with all information in the noise, using a universal model. We can easily show that if such models exist, we can choose our numbering so that they are always chosen.
\begin{restatable}[underfitting]{lemma}{underfitting}
Let $\M$ be a model class containing model $u$, called a \emph{universal model} with $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$. Then, there exist numberings such that $\s^\M$ is bounded.
\end{restatable}
\noindent Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\P$). This eliminates the universal model, but as we shall see, it does not solve the problem for most data. The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}. The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.

@@

We have seen that some variants of sophistication sidestep the simplest examples of underfitting by using total models. Does this ensure that these variants are not bounded? A well-known result in algorithmic statistics shows that there exist strings for which no finite set is as good a model as the singleton set. In this section we will focus on the variants $S^\T$ and $S^T_K$ \cite{vitanyi2004meaningful}, as those are the variants most resistant to proofs of non-invariance.

\begin{restatable}{lemma}{nonstochastic}
There are infinitely  $x$ such that $S^\F_K(x) \geq K(x)/5$. \label{lemma:structure-function-is-not-bounded}
\end{restatable}

\begin{restatable}{lemma}{vitanyi}
Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$. Then there exists a finite set $S$ containing $x$ and a constant $c$, such that $K(S) \leq K(f) + K(|d|) + c$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
\end{restatable}
\noindent This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}. We use this lemma to transport the result of Theorem~\ref{lemma:structure-function-is-not-bounded} to the model class of total functions:

@@

The \emph{$c$-depth} $d^\M_c(x)$ of a string is $d^\M_c(x) = \min \left\{t : C^\M_t(x) - C^\M(x) \leq c \right\}$.
\end{definition}
Deep strings are those that can only be optimally compressed with a great investment of time. We note that it is exceedingly unlikely that a deep string is sampled from a shallow distribution \cite{bloem2014safe,bennett1988logical}.

@@

\begin{lemma}[Invariance of function complexity]
Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be any two acceptable numberings. There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function such that $\psi_i$ to $\phi_{g(i)}$.
\begin{align*}
K_\phi(f) &= \min\left\{ K_\phi(i) : \phi_i= f\right\} 
\geq \min\left\{ K_\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K_\psi(i) : \psi_{g(i)}= f\right\} - c
\geq \min\left\{ K_\psi(i) : \psi_i= f\right\} - c = K_\psi(f).
\end{align*}
The opposite inequality can be achieved by reversing $\phi$ and $\psi$. 
\end{proof}
This lemma shows that the complexity of a function does not depend on the choice of
enumeration by more than a constant term.

\subsection{Existence of faithful numberings}

@@

Let $\num{\phi}$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.

>>>>

\sdr{Course sophistication handled sufficiently?}
\sdr{Do we really need the $\gamma$-invariance? Or can $\gamma$ be 1?}
\sdr{We need the definitions! But maybe we can simplify, do we really need the candidate set?}

Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring. There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity. For example, a sequence of a million coin flips will almost certainly have maximal Kolmogoroc complexity, even though there is nothing complex about flipping a lot of coins---it's just hard work. For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data. This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed. In this paper, we investigate three serious problems with these definitions, and show why they are unsatisfactory. We believe that the problems are fundamental, and provide arguments why such attempts to define sophistication must fail.

The Kolmogorov complexity of a binary string $x$, denoted $K(x)$, is (roughly) the length of the shortest computer program that will output $x$. This length depends on which programming language is used, but, as shown by a core result known as the Invariance Theorem, the language affects the result only by a constant, independent of $x$. This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings. We can see it as \emph{a property of the data}.

A robust definition of sophistication $S(x)$ in the spirit of Kolmogorov complexity should have similar robustness guarantees. We are interested in a measure satisfying the following requirements:

@@

Underfitting (Section~\ref{section:underfitting}) occurs when the selected model is simple, but fails to represent some of the structure present in the data. This is a problem for sophistication because the considered models can be so powerful. In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language. Such \emph{universal models} can be described in a constant amount of bits, and  represent the data with an input of the length of the Kolmogorov complexity, essentially encoding all information as noise. If complex models are penalized, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data. The usual workaround is to restrict the set of allowed models to exclude the universal ones. For example, the model class may be restricted to the set of all total recursive functions. While excludes universal models, it is questionable whether it adequately solves the problem of underfitting in general.

@@

\section{Preliminaries}
Let $\B = \{0,1\}^*$. We deal with computable partial functions $\phi: \B \to \B$, which we will call \emph{models}. A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix free set. A function $f$ is \emph{total} if $D_f = \B$. A prefix-free function is \emph{complete} if every infinite binary string has a member of $D_f$ as a prefix. We fix a prefix code with $|\bar{x}| \le|x|+2\log(|x|+2)|$ (eg. \cite[Section~1.4]{li1993introduction}).

\sdr{I think this completeness vs totalness should be explained much better, its important because without it the differences between the settings cannot be explained}

\begin{definition}[Numberings]

\begin{definition}[Prefix encoding function]
  A prefix encoding function $f:\B\to\B$ is an injective
  function that maps the binary strings to a prefix-free set. There
  are prefix encoding functions $g$ with the property that
  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
  denote it using a bar, $\br{x}=g(x)$.
\end{definition}

\sdr{Clearly explain what it means if a function function is total, complete, prefix. Do we need to fuss about equating the numbers and binary strings? Perhaps it's better to do the conversions explicitly in this paper}

We will generalize the main proposals for sophistication into a single notation. We start with a class of models, we consider all two-part descriptions of the data using these models, we select the set of representations with the shortest code-lengths and we select the smallest model from this set. The size of the smallest model represents the sophistication of the data.

A numbering is an enumeration of of the partial computable functions, denoted as $\psi_1, \psi_2, \ldots$ or simply $\psi$. We fix one canonical effective numbering $\phi$. We call a numbering $\psi$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ with $\forall: i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{a(i)}$.
\end{definition}

\begin{definition}[Complexity]
Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions.\sdr{Cite textbook that this is possible} We define the plain and prefix Kolmogorov complexity as $C(x)=\min\{|\bar\imath y|:\phi_i(y)=x\}$ and $K(x)=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}$ respectively.
\label{definition:complexity}
\end{definition}
\sdr{Added back, liked it} This definition of $K$ is uncommon. It is equivalent to the standard definition in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).

\begin{definition}[Model classes] A \emph{model class} is a set of models. We distinguish five important model classes: the partial (computable) functions ($\C$), the prefix functions ($\K \subset \C$), the total functions ($\T \subset \C$), the complete functions ($\P \subset \K$) and the finite sets ($\F \subset \P$, encoded as a prefix-free function mapping all strings of length $\lceil\log S\rceil$ to members of $S$).
\end{definition}

\sdr{We used to define: A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model and $m(y)=x$. I don't think we can change this. Otherwise an element of $\F$ an index is not a representation.}

\begin{definition}[Representations]
A pair $(i,y)$ is a \emph{representation} for $x$ if $\phi_i(y)=x$.
Define the following code-length functions for the representations:\\
\-\hspace{1cm}$L(m)$ is the length of a self-delimiting description of the model $m$.\\ 
\-\hspace{1cm}$L^m(x)=L(m)+\min\{|y|:m(y)=x\}$\\
\-\hspace{1cm}$L^\M(x)=\min\{L^m(x):m\in\M\}$
>>>>>>> 3ebd420d297744b1c9bdacf2ebd8051f62dcfd13
\end{definition}
We can now define sophistication generically:
\begin{definition}[Sophistication]
The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
$C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}$
with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $L^\M(x)$. The \emph{sophistication} is \\ 
\-\hspace{1cm}$\s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}$\label{definition:sophistication}
\end{definition}
Finally, we generalize the notion of invariance to allow for less strict versions. 
\begin{definition}[$\gamma-invariance$]
Let $\gamma:\N\to \N$ be some nondecreasing function. Two functions $f$ and $g$ are $\gamma$-equivalent if $|f(x) - g(x)| \in O(\gamma(\max(f(x), g(x)))$. We say that a function $\s_\phi(x)$ is \emph{$\gamma$-invariant} to the choice of numbering if $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$. 
\end{definition}
\begin{lemma}\label{lem:broken}
If $\s_\phi(x)$ is $\gamma$-equivalent to either $0$ or to $K(x)$ for \emph{some} numbering $\phi$, then either this equivalence exists for all other numberings as well, or $\s_\phi(x)$ is not $\gamma$-invariant.
\end{lemma}

\section{Inefficient indices}
\label{section:indices}

We start from a simple intuition: the definition of $K$ already uses a description of the data in two parts; a computable function and an input. Can we assume that, when such a description is optimal, it automatically separates the structural and random information? As we will show, the representation of functions plays crucial fundamental.

We will first illustrate what happens if any acceptable numbering is allowed:
\begin{restatable}{lemma}{ineffindices}
Any sophistication $\s_\phi(x)$ that returns the length of some index $i$ such that $\phi_i(y) = x$ for some $y$, cannot be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
 \end{restatable}
\sdr{Perhaps this proof should be here? It is informative}
\noindent In the proof of this lemma (Section~\ref{section:appendix-inefficient-indices}), we deliberately used ridiculously inefficient index functions. We could use a more efficient index, but most definitions of sophistication do not mention any such restrictions. As we shall see, any inefficiency in the index will lead to a lack of invariance in sophistication.

@@

\begin{definition}[\cite{grunwald2004shannon,vitanyi2004meaningful}]
The complexity of a partial recursive function is defined by $C_\phi(f) = \min\{C(i):\phi_i=f\}$ and $K_\phi(f) = \min\{K(i):\phi_i=f\}$.\label{definition:model-complexity}
\end{definition}
We show in the appendix (Lemma~\ref{lemma:invariance}) that these are invariant, ie. complexity is a property of a function. There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature.
\begin{enumerate}
  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used. Specifically, that for every function $f$ there is an $i$ with $\phi_i=f$ and $|i|=C(f)$. We construct such an index below. This approach is taken by Adriaans \cite{adriaans2012facticity}.
  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use $K$ to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data. With this approach, the total code length is $K(\phi_i)+|y|$, with $\phi_i(y)=x$. In this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints. This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
\end{enumerate}
To show that the two perspectives are equivalent, we must show that there exists an acceptable numbering which is efficient enough:
\begin{definition}[Faithful Numbering]\label{def:faithful}
  A numbering $\psi$ is \emph{faithful} if
  for all $i$ there is a $j$ such that $\psi_i=\psi_j$ and
  $|j|\le C(\psi_j)+c$, for some constant $c$.
\end{definition}
This definition is a refinement of \cite[Definition~10]{adriaans2012facticity}. We show in the appendix (Lemma~\ref{lemma:faithful-numberings}) that such numberings exist. Unfortunately, even with a faithful numbering, we can fail:
\begin{restatable}{lemma}{ineffprefix}
Let $\phi$ be faithful and let $L(\phi_i) = |\bar\imath|$, for some prefix encoding function. If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $\s^\C$ is bounded for all numberings.\label{lemma:prefix-inefficiency}
\end{restatable}
\noindent This result applies to \cite{adriaans2012facticity}. To avoid the problem, the model $\phi_i$ should be encoded in $K(\phi_i)$ bits, relying on $K$'s own prefix encoding to delimit the model. 
\sdr{The above is incomprehensible. Explain better why this can't be fixed; point out that it provides a slight imbalance in favour of putting stuff in the data part, which, if the model class is $\C$, already to underfitting. Prime the reader to think in terms of balance. Refer to the underfitting section.}
\sdr{We should point out from the start that the difference in cost for the model $|\bar\imath|$ instead of $K(i)$ is unbounded, so even a faithful model index is still not an optimal description, and the gap is a direct consequence of the definition of Kolmogorov complexity. And that this slight imbalance can already lead to underfitting. We can then say that the model does not occur if we use the other two-part coding approach, which is why we'll adopt that other perspective in the rest of the paper.} In order to circumvent the problem, the representation should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.
Note that the condition for this lemma likely applies to all possible computable  prefix encoding functions (see Section~\ref{section:prefix-inefficiency}). Thus, even if we have a faithful numbering, we can only use it if we don't count the cost of delimiting the model from the input (as in \cite{koppelSoph1988}). In the following, we will take the second perspective: the representations are two-part descriptions $(i, y)$ such that $\phi_i(y)=x$, with codelength $K(\phi_i) + |y|$.

\sdr{Moved this back here, the conjecture needs to be in the main text}
Lemma~\ref{lemma:prefix-inefficiency} only holds if the prefix function $L(i)$ has the property that $\min\{L(i)-K(i):i\ge i_0\}$ is unbounded. We conjecture that this is the case for any computable function.

\section{Overfitting: the singletons}
\label{section:overfitting}
We now come to the main issue plaguing sophistication. \sdr{Beetje bombastisch. En hierna zou ik zeggen, we first consider the simplest, most natural setting, \dots} If we use model class $\K$, the candidate set always contains two representations that do not provide a good separation. The first is a \emph{universal model}: a basic universal Turing machine. The second is a \emph{singleton model} a model which returns $x$ for any input. As we will show, in many settings, we can ensure that one of them is the \emph{only} model in the candidate set. The following lemma generalizes the principle:

\begin{restatable}{lemma}{coolone}
\label{lemma:thecoolone}
  Let $\psi$ be any acceptable enumeration of the partial recursive functions.
  Let $\M$ be any model class, let $\X$ be any set of binary sequences and let $D:\B\to\N$ be a computable decoding function with a prefix-free domain that maps function descriptions to their indices in $\psi$, i.e. if $f=\psi_{D(p)}$ then $p$ is a $D$-description of $f$. Let $\M'=\{\psi_i:i\in\tn{range}(D)\}$. Further assume there is a constant $c$ such that:\\
\-\hspace{1cm}(1)$\forall_{f\in\M'}:\min\{|p|:\psi_{D(p)}=f\}\le K^\psi(f)+c$\\
\-\hspace{1cm}(2)$\forall_{x\in\X}:L^{\M',\psi}(x)-L^{\M,\psi}(x)\le c$.\\
Then there is an enumeration $\phi$ of the partial recursive functions such that $S^\phi_{\M}(x) = |\bar 0|+S^\psi_{\M'}(x)$ for all $x\in\X$.
\end{restatable}
\noindent We can now show numberings for which the singletons are always selected:

\begin{restatable}[Overfitting]{lemma}{overfitting}
Let $\M \subseteq \K$ be a model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$. Then there is an enumeration $\phi$ of the prefix partial recursive functions, and a constant $c$, such that
\[
K(x)-S^{\M,\phi}(x)\le c
\]
for all $x\in\X$.
\end{restatable}
\noindent In short, this lemma tells us that there are numberings for which the singleton models always determine the sophistication. This means that there is a numbering $\phi$ such that the sophistication is equivalent to Kolmogorov complexity, which implies that this sophistication is unsound (by Lemma~\ref{lem:broken}).

Many variants in the literature (including the structure function) are susceptible to this problem  \sdr{Don't say ``many'', just list explicitly} \cite{cover1985kolmogorov,gacs2001algorithmic,vereshchagin2013algorithmic}. In \cite{vitanyi2004meaningful,adriaans2012facticity} this issue is avoided by not using self-delimiting representations.
\sdr{Prep the balance idea so that this is understandable. We need it}  Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically not optimally efficient. While this dodges the issue of the singletons, it does not show that $S$ is invariant.

@@

\section{Underfitting: the universal model}
\label{section:underfitting}
On the other end of the spectrum, there exist representations with all information in the noise, using a universal model. We can easily show that if such models exist, we can choose our numbering so that they are always chosen.
\begin{restatable}[underfitting]{lemma}{underfitting}
Let $\M$ be a model class containing a universal model $u$, with the property that $\exists c \forall m \in \M : L^u(x) \leq L^m(x) + c \p$. Then, there exist numberings such that $\s^\M$ is bounded.
\end{restatable}
\noindent Most sophistication measures avoid the underfitting problem by limiting their model class, usually to either finite sets ($\F$) or total functions ($\T$ or $\P$). This eliminates the universal model, but as we shall see, it does not solve the problem for most data. The variants that are susceptible to the this type of underfitting are facticity \cite{adriaans2012facticity} and some versions of effective complexity \cite{gellmann1996information}. The latter aims to bypass the problem by selecting carefully from the candidate set, but as we have shown, for some numberings the candidate set contains only the universal model.
\sdr{Explain this is actually a widely recognised problem}

@@

We have seen that some variants of sophistication sidestep the simplest examples of underfitting by using total models. Does this ensure that these variants are not bounded? A well-known result in algorithmic statistics shows that there exist strings for which no finite set is as good a model as the singleton set. In this section we will focus on the variants $S^\T$ and $S^T_K$ \cite{vitanyi2004meaningful}, as those are the variants most resistant to proofs of non-invariance.
\sdr{Do I know this notation? Also, I wanna emphasize that the structure function underfits in this way.}
\sdr{Combine this lemma with the one after next}
\begin{restatable}{lemma}{nonstochastic}
There are infinitely  $x$ such that $S^\F_K(x) \geq K(x)/5$. \label{lemma:structure-function-is-not-bounded}
\end{restatable}

\sdr{Move this lemma to the appendix altogether}
\begin{restatable}{lemma}{vitanyi}
Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$. Then there exists a finite set $S$ containing $x$ and a constant $c$, such that $K(S) \leq K(f) + K(|d|) + c$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
\end{restatable}
\noindent This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}. We use this lemma to transport the result of Theorem~\ref{lemma:structure-function-is-not-bounded} to the model class of total functions:

@@

The \emph{$c$-depth} $d^\M_c(x)$ of a string is $d^\M_c(x) = \min \left\{t : C^\M_t(x) - C^\M(x) \leq c \right\}$.
\end{definition}

Deep strings are those that can only be optimally compressed with a great investment of time. We note that it is exceedingly unlikely that a deep string is sampled from a shallow distribution \cite{bloem2014safe,bennett1988logical}.

@@

\begin{lemma}[Invariance of function complexity]
Let $\phi$ and $\psi$ be any two acceptable numberings. There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function such that $\psi_i=\phi_{g(i)}$.
\begin{align*}
K_\phi(f) &= \min\left\{ K_\phi(i) : \phi_i= f\right\} 
\geq \min\left\{ K_\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K_\psi(i) : \psi_{g(i)}= f\right\} - c
\geq \min\left\{ K_\psi(i) : \psi_i= f\right\} - c' = K_\psi(f).
\end{align*}
The opposite inequality can be achieved by reversing $\phi$ and $\psi$. 
\end{proof}
This lemma shows that the complexity of a function does not depend on the choice of
enumeration by more than a constant term.

@@

Let $\phi$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.
